---
title: "Cox Model"
author: "Youlan Shen"
output: github_document
---

```{r setup, include=FALSE}
library("survival")
library(StepReg)
library(tidyverse)
library("readxl")
library(ggsurvfit) # for better visualization
library(survminer)
library(dplyr)
data(cancer, package = "survival")
```

## Cox Model Assumption Checking

### checking if the proportional hazard assumption between two treatment groups 

```{r}
# data preprocessing
head(rotterdam)
# liberal definition of rfs (count later deaths)
rotterdam <- rotterdam %>% 
  mutate(rfs = pmax(recur, death),
         rfstime = ifelse(recur == 1, rtime, dtime))
```


```{r}
colon_survfit_log = survfit(Surv(log(rfstime + 1), rfs) ~ hormon, data = rotterdam)
splots <- list()
splots[[2]] <- ggsurvplot(colon_survfit_log, data = rotterdam, fun = "cloglog",
                          risk.table = FALSE, xlab = "log(Time)", 
                          ggtheme = theme_minimal(), xlim = c(3,10))
splots[[2]]$plot + labs(title = "Log of Negative Log of Estimated Survival Function-hormon")

colon_survfit_log_2 = survfit(Surv(log(rfstime + 1), rfs) ~ grade, data = rotterdam)
splots_2 <- list()
splots_2[[2]] <- ggsurvplot(colon_survfit_log_2, data = rotterdam, fun = "cloglog",
                          risk.table = FALSE, xlab = "log(Time)", 
                          ggtheme = theme_minimal(), xlim = c(3,10))
splots_2[[2]]$plot + labs(title = "Log of Negative Log of Estimated Survival Function-grade")
```

From the log-log graph, the lines for two treatment groups are not proportional to each other, therefore, the proportional hazard assumption violates.

The line is close to a straight line with slope equal to 1, we could consider using exponential distribution.

### trying stratified groups on size groups/age groups

```{r}

```

### check continuous variables interaction with time

```{r}
# fit a entire cox model, find significant variables
# er and meno, seems less significant
fit1 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo, data = rotterdam)

# change the data to variables with interaction with time
# nodes become no significant, pgr and er more significant
# except nodes, all other continuous variables have significant interaction with time
rotterdam <- rotterdam %>% 
  mutate(
    aget = age * log(rfstime),
    gradet = grade * log(rfstime),
    nodest = nodes * log(rfstime),
    pgrt = pgr * log(rfstime), 
    ert = er * log(rfstime)
)
fit2 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo + aget + gradet + nodest + pgrt + ert, data = rotterdam)
```

### model fitting and plots

```{r}
# stepwise model
model3 <- stepwiseCox(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo + aget + gradet + nodest + pgrt + ert, data = rotterdam, 
            selection = "bidirection", select = "SL", method = "breslow", 
            sle = 0.25, sls = 0.15)
model3$`Selected Varaibles`
fit3 <- coxph(Surv(rfstime, rfs) ~ gradet + grade	+ nodest + hormon + age	+ aget + size +
                meno + ert + pgrt + pgr + er, data = rotterdam)
```

### observed and fitted plot/ fitted survival plot?

```{r}

```

### model fit results

```{r}
summary(fit1)
summary(fit2)
summary(fit3)
```

