---
title: "Cox Model"
author: "Youlan Shen"
output: github_document
---

```{r setup, include=FALSE}
library("survival")
library(StepReg)
library(tidyverse)
library("readxl")
library(ggsurvfit) # for better visualization
library(survminer)
library(dplyr)
library(patchwork)
library(MASS)
data(cancer, package = "survival")
```

## Cox Model Assumption Checking

### checking if the proportional hazard assumption between two treatment groups 

```{r}
# data preprocessing
head(rotterdam)
# liberal definition of rfs (count later deaths)
rotterdam <- rotterdam %>% 
  mutate(rfs = pmax(recur, death),
         rfstime = ifelse(recur == 1, rtime, dtime),
         grade = factor(grade)) 
```


```{r}
colon_survfit_log = survfit(Surv(log(rfstime + 1), rfs) ~ hormon, data = rotterdam)
splots <- list()
splots[[2]] <- ggsurvplot(colon_survfit_log, data = rotterdam, fun = "cloglog",
                          risk.table = FALSE, xlab = "log(Time)", 
                          ggtheme = theme_minimal(), xlim = c(3,10))
splots[[2]]$plot + labs(title = "Log of Negative Log of Estimated Survival Function")
```

From the log-log graph, the lines for two treatment groups are not proportional to each other, therefore, the proportional hazard assumption violates.

The line is close to a straight line with slope equal to 1, we could consider using exponential distribution.



### trying stratified groups on size groups/age groups

```{r}
size <- rotterdam %>% dplyr::select(size) %>% unique() %>% pull()
plot <- list()
for (i in 1:length(size)) {
  data <- rotterdam[rotterdam$size == size[i], ]
  colon_survfit_log = survfit(Surv(log(rfstime + 1), rfs) ~ hormon, data = data)
  splots <- ggsurvplot(colon_survfit_log, data = data, fun = "cloglog",
                          risk.table = FALSE, xlab = "log(Time)", 
                          ggtheme = theme_minimal(), xlim = c(3,10)) + labs(title = paste("Log(-Log(S(t))) of in Size Group", size[i]))
  plot[[i]] <- splots
}
arrange_ggsurvplots(plot, print = TRUE,
  ncol = 2, nrow = 2, risk.table.height = 0.4)

```

### check continuous variables interaction with time (not included since assumption is not violated)

```{r}
# fit a entire cox model, find significant variables
# er and meno, seems less significant
fit1 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo, data = rotterdam)

# change the data to variables with interaction with time
# nodes become no significant, pgr and er more significant
# except nodes, all other continuous variables have significant interaction with time
rotterdam <- rotterdam %>% 
  mutate(
    aget = age * log(rfstime),
    nodest = nodes * log(rfstime),
    pgrt = pgr * log(rfstime), 
    ert = er * log(rfstime)
)
fit2 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo + aget + nodest + pgrt + ert, data = rotterdam)
```

### model fitting and plots

```{r}
# stepwise model
model3 <- stepwiseCox(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo + aget + nodest + pgrt + ert, data = rotterdam, 
            selection = "bidirection", select = "SL", method = "breslow", 
            sle = 0.25, sls = 0.15)
model3$`Selected Varaibles`
fit3 <- coxph(Surv(rfstime, rfs) ~ nodes + nodest + age	+ aget + meno + er + 
                ert + hormon + grade + size, data = rotterdam)
```

```{r}
fit4 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo, data = rotterdam)
stepwise_model <- stepAIC(fit4, direction = "both", trace = FALSE)
```

### observed and fitted plot/ fitted survival plot?

```{r}

```

### model fit results

```{r}
summary(fit1)
summary(fit2)
summary(fit3)
summary(fit4)
summary(stepwise_model)
```

