---
title: "Log-rank Test"
author: "Zhezheng Jin"
date: "2023-11-30"
output: github_document
---

```{r setup}
library(survival)
library(nph)
```

## Log-rank test (combine both as event) comparing hormonal treatment
```{r}
rfs  <- pmax(rotterdam$recur, rotterdam$death)
rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))
surv_obj1 <- Surv(time = rfstime, event = rfs)
log_rank_test_hormon1 <- survdiff(surv_obj1 ~ hormon, data = rotterdam)

print(log_rank_test_hormon1)
```

The p-value of 0.0003 suggests that there is a statistically significant difference in survival between the two hormone treatment groups.
The very low p-value (much less than 0.05) leads us to reject the null hypothesis of no difference in survival between the two groups.
The result indicates that hormone treatment status (hormon) is strongly associated with differences in survival outcomes in the dataset, when considering both recurrence and death as the events of interest. Specifically, the observed number of events in each group significantly deviates from what would be expected if there were no difference in survival between the groups.

## Stratified log-rank test (combine both as event) comparing hormonal treatment stratified by chemotherapy
```{r}
stratified_log_rank_test <- survdiff(surv_obj3 ~ hormon + strata(chemo), data = rotterdam)

print(stratified_log_rank_test)
```

Stratification: The test was stratified by the chemotherapy, meaning the analysis was conducted separately within each chemotherapy status group, but the overall test combines these analyses.

The p-value of 0.0001 suggests that there is a statistically significant difference in survival between the two hormone treatment groups when stratified by chemotherapy status.
The very low p-value (much less than 0.05) leads us to reject the null hypothesis of no difference in survival between the two groups within each stratum of chemotherapy status.
This result indicates that hormone treatment status (hormon) is strongly associated with differences in the combined outcomes of recurrence and death, and this association holds true even when accounting for whether patients received chemotherapy or not.

## Gehan's Wilcoxon Test
```{r}
gehan_test <- survdiff(surv_obj1 ~ hormon, data = rotterdam, rho = 1)
print(gehan_test)
```

survdiff(formula = surv_obj1 ~ hormon, data = rotterdam, rho = 1)

            N Observed Expected (O-E)^2/E (O-E)^2/V
hormon=0 2643     1057     1087     0.864      11.2
hormon=1  339      151      121     7.795      11.2

 Chisq= 11.2  on 1 degrees of freedom, p= 8e-04 

## Peto & Prentice Test
```{r}
peto_test <- survdiff(surv_obj1 ~ hormon, data = rotterdam, rho = 0)
print(peto_test)
```

survdiff(formula = surv_obj1 ~ hormon, data = rotterdam, rho = 0)

            N Observed Expected (O-E)^2/E (O-E)^2/V
hormon=0 2643     1507     1551      1.25      13.4
hormon=1  339      206      162     12.02      13.4

 Chisq= 13.4  on 1 degrees of freedom, p= 3e-04 

## Fleming-Harrington Test
```{r}
# p=0, q=1
fh_test2 <- logrank.test(rotterdam$rfstime, rotterdam$rfs, rotterdam$hormon, rho = 0,  gamma = 1)
print(fh_test2)

# p=1, q=0
fh_test1 <- logrank.test(rotterdam$rfstime, rotterdam$rfs, rotterdam$hormon, rho = 1,  gamma = 0)
print(fh_test1)
```

logrank.test(time = rotterdam$rfstime, event = rotterdam$rfs, 
    group = rotterdam$hormon, rho = 0, gamma = 1)

     N Observed Expected (O-E)^2/E (O-E)^2/V
1 2643     1507     1551      1.25       145
2  339      206      162     12.02       145

 Chisq= 13.6  on 1 degrees of freedom, p= 2e-04
 rho   =  0 gamma =  1 
 
logrank.test(time = rotterdam$rfstime, event = rotterdam$rfs, 
    group = rotterdam$hormon, rho = 1, gamma = 0)

     N Observed Expected (O-E)^2/E (O-E)^2/V
1 2643     1507     1551      1.25      23.2
2  339      206      162     12.02      23.2

 Chisq= 11.2  on 1 degrees of freedom, p= 8e-04
 rho   =  1 gamma =  0 

## Tarone-Ware Test
```{r}
tarone_ware_test <- survdiff(surv_obj1 ~ hormon, data = rotterdam, rho = 0.5)
print(tarone_ware_test)
```

survdiff(formula = surv_obj1 ~ hormon, data = rotterdam, rho = 0.5)

            N Observed Expected (O-E)^2/E (O-E)^2/V
hormon=0 2643     1251     1287      1.04      12.4
hormon=1  339      175      139      9.63      12.4

 Chisq= 12.4  on 1 degrees of freedom, p= 4e-04 



