---
title: "Partial likelihood Test"
author: "Zhezheng Jin"
date: "2023-11-30"
output: github_document
---

```{r setup}
library(survival)
```

## Partial likelihood Test for hormonal treatment

```{r}
rfs  <- pmax(rotterdam$recur, rotterdam$death)
rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))

# Fit the full model
full_model <- coxph(Surv(rfstime, rfs) ~ age + size + grade + nodes + pgr + er + hormon, data = rotterdam)

# Fit the reduced model (without hormon)
reduced_model <- coxph(Surv(rfstime, rfs) ~ age + size + grade + nodes + pgr + er, data = rotterdam)

# Conduct the likelihood ratio test
test <- anova(reduced_model, full_model, test="LRT")

# View the results
print(test)
```

The partial likelihood ratio test comparing two Cox proportional hazards models: one without the 'hormon' variable (Model 1) and one with it (Model 2). Based on the log-likelihood values, Model 2 (including the 'hormon' variable) has a slightly higher log-likelihood (-12567) compared to Model 1 (-12569), suggesting a marginally better fit.
The p-value is 0.112. This p-value tests the null hypothesis that adding 'hormon' to the model does not significantly improve the model's fit.