---
title: "Survival Analysis of Breast Cancer Patients: Investigating Hormone Treatment Effects on Survival Rates Using the Rotterdam Dataset"
author: "Zhezheng Jin, Mingkai Chen, Youlan Shen, Hongru Tang "
date: "2023-12-11"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \usepackage{algorithm}
- \usepackage{algpseudocode}
- \usepackage{amsthm}
---

```{r setup, include=FALSE, eval = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
library(knitr)
library(tibble)
library(survival)
library(dplyr)
library(ggplot2)
library(survminer)
library(gtsummary)
library(kableExtra)
library(gridExtra)
library(survival)
library(MASS)
library(ggplot2)
library(tidyverse)
library(survminer)
library(survival)
library(nph)
library("survival")
library(StepReg)
library(tidyverse)
library("readxl")
library(ggsurvfit) # for better visualization
library(survminer)
library(dplyr)
library(patchwork)
library(MASS)
```

## discription in two treatment

```{r eval=FALSE}
head(rotterdam)
Rotterdam=data.frame(survival::rotterdam)
colSums(is.na(Rotterdam))
stats <- Rotterdam %>% 
  mutate(
    treat = case_when(
      hormon == 0 ~ "Treat = 0",
      hormon == 1 ~ "Treat = 1"))%>% 
  dplyr::select(-pid, -hormon, -rtime, -recur, -dtime, -death)

stats <- stats %>% 
  bind_rows(stats %>% 
              mutate(treat = "Overall"))
stats %>%
  tbl_summary(by = treat) %>% 
   modify_caption("**People Characteristics for Hormon Treatment**") %>%
  bold_labels()
```

## summary of variables

```{r eval=FALSE}
stats2 <- Rotterdam %>% 
  mutate(
    treat = case_when(
      chemo == 0 ~ "Treat = 0",
      chemo == 1 ~ "Treat = 1"))%>% 
  dplyr::select(-pid, -chemo, -rtime, -recur, -dtime, -death)
stats2 <- stats2 %>% 
  bind_rows(stats2 %>% 
              mutate(treat = "Overall"))
stats2 %>%
  tbl_summary(by = treat) %>% 
   modify_caption("**People Characteristics for Chemo Treatment**") %>%
  bold_labels()
```

## KM estimation

```{r eval=FALSE}
status  <- pmax(rotterdam$recur, rotterdam$death)
rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))

surv_obj <- Surv(time = rfstime, event = status)

km_fit1 <- survfit(surv_obj ~ 1)
print(km_fit1)

ggsurvplot(km_fit1, data=  rotterdam,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark2",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Time",              
           ylab = "Survival Probability") 
```


```{r eval=FALSE}
km_fit2 <- survfit(surv_obj ~ Rotterdam$hormon)
print(km_fit2)


ggsurvplot(km_fit2, data=  rotterdam,                  
           pval = TRUE,                
           conf.int = TRUE, 
           risk.table = TRUE,
           ggtheme = theme_minimal(),  
           palette = "Dark2",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Time",              
           ylab = "Survival Probability") 
ggsave(file = "K-M-estimation-part_files/KM1.jpg")
```


```{r eval=FALSE}
Rotterdam$status  <- pmax(rotterdam$recur, rotterdam$death)
Rotterdam$rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))

rotterdam_0 <- Rotterdam%>%filter(grade == "2" & size == "<=20" )
rotterdam_1 <- Rotterdam%>%filter(grade == "3" & size == ">50" & nodes >=5)
rotterdam_2 <- Rotterdam%>%filter(size == ">50")
km_fit3 <- survfit(Surv(rfstime, status) ~ rotterdam_0$hormon,data = rotterdam_0)
print(km_fit3)
ggsurvplot(km_fit3, data=  rotterdam_0,                  
           conf.int = TRUE,  
           risk.table = TRUE,
           ggtheme = theme_minimal(),  
           palette = "Dark2",         
           title  = "Kaplan-Meier Curve for Hormon = 0", 
           xlab = "Time",              
           ylab = "Survival Probability") 
ggsave(file = "K-M-estimation-part_files/KM2.jpg")
```


## Log-rank test (combine both as event) comparing hormonal treatment
```{r eval=FALSE}
rotterdam$rfs  <- pmax(rotterdam$recur, rotterdam$death)
rotterdam$rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))
surv_obj1 <- Surv(time = rfstime, event = rfs)
log_rank_test_hormon1 <- survdiff(surv_obj1 ~ hormon, data = rotterdam)

print(log_rank_test_hormon1)
```

The p-value of 0.0003 suggests that there is a statistically significant difference in survival between the two hormone treatment groups.
The very low p-value (much less than 0.05) leads us to reject the null hypothesis of no difference in survival between the two groups.
The result indicates that hormone treatment status (hormon) is strongly associated with differences in survival outcomes in the dataset, when considering both recurrence and death as the events of interest. Specifically, the observed number of events in each group significantly deviates from what would be expected if there were no difference in survival between the groups.

## Stratified log-rank test (combine both as event) comparing hormonal treatment stratified by chemotherapy
```{r eval=FALSE}
stratified_log_rank_test <- survdiff(surv_obj3 ~ hormon + strata(chemo), data = rotterdam)

print(stratified_log_rank_test)
```

Stratification: The test was stratified by the chemotherapy, meaning the analysis was conducted separately within each chemotherapy status group, but the overall test combines these analyses.

The p-value of 0.0001 suggests that there is a statistically significant difference in survival between the two hormone treatment groups when stratified by chemotherapy status.
The very low p-value (much less than 0.05) leads us to reject the null hypothesis of no difference in survival between the two groups within each stratum of chemotherapy status.
This result indicates that hormone treatment status (hormon) is strongly associated with differences in the combined outcomes of recurrence and death, and this association holds true even when accounting for whether patients received chemotherapy or not.

## Gehan's Wilcoxon Test
```{r eval=FALSE}
gehan_test <- survdiff(surv_obj1 ~ hormon, data = rotterdam, rho = 1)
print(gehan_test)
```



## Peto & Prentice Test
```{r eval=FALSE}
peto_test <- survdiff(surv_obj1 ~ hormon, data = rotterdam, rho = 0)
print(peto_test)
```



## Fleming-Harrington Test
```{r eval=FALSE}
# p=0, q=1
fh_test2 <- logrank.test(rotterdam$rfstime, rotterdam$rfs, rotterdam$hormon, rho = 0,  gamma = 1)
print(fh_test2)

# p=1, q=0
fh_test1 <- logrank.test(rotterdam$rfstime, rotterdam$rfs, rotterdam$hormon, rho = 1,  gamma = 0)
print(fh_test1)
```



## Tarone-Ware Test
```{r eval=FALSE}
tarone_ware_test <- survdiff(surv_obj1 ~ hormon, data = rotterdam, rho = 0.5)
print(tarone_ware_test)
```


```{r eval=FALSE}
rfs  <- pmax(rotterdam$recur, rotterdam$death)
rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))
rotterdam$tstart = c(rep(0,2982))
surv_object <- with(rotterdam, Surv(tstart, rfstime, rfs)) #with consideration of time-varying covariates
```


```{r eval=FALSE}
rotterdam$rfs  <- pmax(rotterdam$recur, rotterdam$death)
rotterdam$rfstime <- with(rotterdam, ifelse(recur==1, rtime, dtime))

hormon1 = subset(rotterdam,hormon == 1)
hormon0 = subset(rotterdam, hormon == 0)
unique_hormon_statuses <- unique(rotterdam$hormon)
# Initialize an empty data frame for plotting
plot_data <- data.frame()

for(status in unique_hormon_statuses) {
  subset_data <- rotterdam[rotterdam$hormon == status,]
  fit <- survfit(Surv(event = rfs,time = rfstime) ~ 1, data = subset_data)
  
  temp_data <- data.frame(time = fit$time, surv = fit$surv, hormon_status = status)
  temp_data$log_surv = log(temp_data$surv)
  # Combine with the main plot data
  plot_data <- rbind(plot_data, temp_data)

} 

ggplot(plot_data, aes(x=time, y=log_surv, color=factor(hormon_status))) +
geom_line() +
geom_point()+
labs(title = "Log(Survival) vs. Time for Different hormon Statuses",
x = "Time",
y = "Log(Survival)",
color = "hormon Status")
```

```{r eval=FALSE}
plot_data = data.frame()
for(status in unique_hormon_statuses) {
subset_data <- rotterdam[rotterdam$hormon == status,]
fit <- survfit(Surv(event = rfs, time = rfstime) ~ 1, data = subset_data)
# Creating a dataframe for this status
temp_data <- data.frame(time = fit$time, surv = fit$surv, hormon_status = status)
temp_data$log_time = log(temp_data$time)
temp_data$log_log_surv = log(-log(temp_data$surv))
# Combine with the main plot data
plot_data <- rbind(plot_data, temp_data)
}
# Plot
ggplot(plot_data, aes(x=log_time, y=log_log_surv, color=factor(hormon_status))) +
geom_line() +
geom_point()+
labs(title = "Log-Log Survival Plot for Different hormon Statuses",
x = "Log(Time)",
y = "Log(-Log(Survival))",
color = "hormon Status")




```


```{r eval=FALSE}
# Set up the plotting area to have 2 rows and 2 columns
par(mfrow = c(2, 2))

# Fit a basic Cox proportional hazards model
cox_model <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er + hormon + chemo, data = rotterdam)
summary(cox_model)
# Checking the Proportional Hazards assumption
cox.zph_model <- cox.zph(cox_model)

# Extract the names of the covariates to use as plot titles
covariate_names <- names(rotterdam[3:11])
cox.zph_model
# Plot Schoenfeld residuals for each covariate with a title
for(i in 1:4){
  plot(cox.zph_model, var = i)
  title(main = paste("Schoenfeld Residuals for", covariate_names[i]))
}
ggsave(file = "cox_ph_files/Schoenfeld.jpg")
```

```{r eval=FALSE}
# Set up the plotting area to have 2 rows and 2 columns
par(mfrow = c(2, 3))

# Fit a basic Cox proportional hazards model
cox_model <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er + hormon + chemo, data = rotterdam)
summary(cox_model)
# Checking the Proportional Hazards assumption
cox.zph_model <- cox.zph(cox_model)

# Extract the names of the covariates to use as plot titles
covariate_names <- names(rotterdam[3:11])

# Plot Schoenfeld residuals for each covariate with a title
for(i in 5:9){
  plot(cox.zph_model, var = i)
  title(main = paste("Schoenfeld Residuals for", covariate_names[i]))
}
ggsave(file = "cox_ph_files/Schoenfeld.jpg")



```


```{r eval=FALSE}
categorical_vars <- c(4, 5, 6, 10, 11)
categorical_names <- c("meno", "size", "grade", "hormon", "chemo")

# Ensure 'rotterdam' is your data frame and 'surv_object' is correctly defined before this loop

for (i in 1:5) {
  col_index <- categorical_vars[i]
  # Fit Kaplan-Meier estimator
  fit <- survfit(surv_object ~ rotterdam[,col_index], data = rotterdam)
  
  # Plot using ggsurvplot
  p <- ggsurvplot(
    fit, 
    data = rotterdam, 
    fun = "cloglog",        # For complementary log-log survival function
    ggtheme = theme_minimal(),
    title = paste("Log-Log plot for", categorical_names[i]),
    xlab = "Log(Time)",
    ylab = "Log(Cumulative hazard)"
  )
  
  # Print the plot
  print(p$plot)
}



```


## Cox Model Assumption Checking

### checking if the proportional hazard assumption between two treatment groups 

```{r eval=FALSE}
# data preprocessing
head(rotterdam)
# liberal definition of rfs (count later deaths)
rotterdam <- rotterdam %>% 
  mutate(rfs = pmax(recur, death),
         rfstime = ifelse(recur == 1, rtime, dtime),
         grade = factor(grade)) 
```


```{r eval=FALSE}
colon_survfit_log = survfit(Surv(log(rfstime + 1), rfs) ~ hormon, data = rotterdam)
splots <- list()
splots[[2]] <- ggsurvplot(colon_survfit_log, data = rotterdam, fun = "cloglog",
                          risk.table = FALSE, xlab = "log(Time)", 
                          ggtheme = theme_minimal(), xlim = c(3,10))
splots[[2]]$plot + labs(title = "Log of Negative Log of Estimated Survival Function")
```

From the log-log graph, the lines for two treatment groups are not proportional to each other, therefore, the proportional hazard assumption violates.

The line is close to a straight line with slope equal to 1, we could consider using exponential distribution.



### trying stratified groups on size groups/age groups

```{r eval=FALSE}
size <- rotterdam %>% dplyr::select(size) %>% unique() %>% pull()
plot <- list()
for (i in 1:length(size)) {
  data <- rotterdam[rotterdam$size == size[i], ]
  colon_survfit_log = survfit(Surv(log(rfstime + 1), rfs) ~ hormon, data = data)
  splots <- ggsurvplot(colon_survfit_log, data = data, fun = "cloglog",
                          risk.table = FALSE, xlab = "log(Time)", 
                          ggtheme = theme_minimal(), xlim = c(3,10)) + labs(title = paste("Log(-Log(S(t))) of in Size Group", size[i]))
  plot[[i]] <- splots
}
arrange_ggsurvplots(plot, print = TRUE,
  ncol = 2, nrow = 2, risk.table.height = 0.4)

```

### check continuous variables interaction with time (not included since assumption is not violated)

```{r}
# fit a entire cox model, find significant variables
# er and meno, seems less significant
fit1 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo, data = rotterdam)

# change the data to variables with interaction with time
# nodes become no significant, pgr and er more significant
# except nodes, all other continuous variables have significant interaction with time
rotterdam <- rotterdam %>% 
  mutate(
    aget = age * log(rfstime),
    nodest = nodes * log(rfstime),
    pgrt = pgr * log(rfstime), 
    ert = er * log(rfstime)
)
fit2 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo + aget + nodest + pgrt + ert, data = rotterdam)
```

### model fitting and plots

```{r eval=FALSE}
# stepwise model
model3 <- stepwiseCox(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo + aget + nodest + pgrt + ert, data = rotterdam, 
            selection = "bidirection", select = "SL", method = "breslow", 
            sle = 0.25, sls = 0.15)
model3$`Selected Varaibles`
fit3 <- coxph(Surv(rfstime, rfs) ~ nodes + nodest + age	+ aget + meno + er + 
                ert + hormon + grade + size, data = rotterdam)
```

```{r eval=FALSE}
fit4 <- coxph(Surv(rfstime, rfs) ~ age + meno + size + grade + nodes + pgr + er +
                hormon + chemo, data = rotterdam)
stepwise_model <- stepAIC(fit4, direction = "both", trace = FALSE)
```


### model fit results

```{r eval=FALSE}
summary(fit1)
summary(fit2)
summary(fit3)
summary(fit4)
summary(stepwise_model)
```




